---
title: "Lab 1"
format: html
editor: visual
---

## Introduction

```{r}
#| label: load-packages
#| message: false
#| echo: false

library(tidyverse)
library(openintro)

```

```{r}
#| label: Introduction-code

data('arbuthnot', package='openintro')

arbuthnot

glimpse(arbuthnot)
```

### Exercise 1

The command to view the counts of the girls baptized will be sum(arbuthnot\$girls) output seen below

```{r}
#| label: Exersise-1-code
sum(arbuthnot$girls)
```

## Data visualization

```{r}
#| label: Data-visualization
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_point()

ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_line()

```

### Exercise 2

Analyzing the previous plots shows us that the amount of baptisms for girls shrunk dramatically between 1640 and 1650. However, in 1660 the amount of girls baptized began to shoot up and had a fairly consistent trend into the early 1700s with a few down years that never met the lows of pre 1660.

## R as a calculator

```{r}
#| label: R-calculator
#| 
arbuthnot$boys + arbuthnot$girls

# |> as a piping tool is just a bit easier to type 
arbuthnot <- arbuthnot |> 
  mutate(total = boys + girls)

ggplot(data = arbuthnot, aes(x = year, y = total)) + 
  geom_line()

5218 / 4683

5218 / (5218 + 4683)

arbuthnot <- arbuthnot |>
  mutate(boy_ratio = boys / total)

ggplot( data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_line()

mean(arbuthnot$boy_ratio)
max(arbuthnot$boy_ratio)
```

### Exercise 3 

When looking at the ratio of boys to girls I can see that it is consistently above the 50 percent mark with the mean being 51.6% and reaching a peak of 53.6%

## More Practice

```{r}
#| label: more-practice
#| 
data('present', package='openintro')

arbuthnot |> 
  summarize(min=min(boys), max=max(boys))

present |> 
  summarize(min=min(year), max=max(year))

years<-present |> 
  summarize(min=min(year), max=max(year))

dim(present)
names(present)
```

### Exercise 4

The years included in the dataset are `{r} years`, the dimensions of the dataset are `{r} dim(present)`, and the names of the columns in the dataset are `{r} names(present)`

### Exercise 5 

```{r}
#| label: exercise-5

present$total <- present$boys + present$girls

ggplot(data = present, aes(x = year, y = total) ) + 
         geom_line()
```

The magnitude is on a completely different scale when comparing the 1600s dataset to the modern day dataset. The modern day dataset is several times larger than the Arbuthnot dataset

### Exercise 6

```{r}
#| label: Exercise-6

present <- present |> 
  mutate( boys_ratio = boys / total )

ggplot( data = present, aes( x = year, y = boys_ratio)) +
  geom_line()
```

The margin in which boys are born more often than girls is much lower, but it seems to show a trend that boys have continued to be born slightly more than girls.

### Exercise 7 

```{r}
#| label: exercise-7

present[  which(present$total == max(present$total)), 'year']


#or to follow the exercise suggestion 

present |> 
  arrange(desc(total))
```

The highest number of births come from the baby boomer generation of 1961
