---
title: "Distribution of Random Variables"
format: pdf
editor: visual
---

# The Normal Distribution 

```{r}
#| title: packages 

library(tidyverse)
library(openintro)

data("fastfood", package='openintro')
head(fastfood)

mcdonalds <- fastfood |>
  filter(restaurant == "Mcdonalds")
dairy_queen <- fastfood |>
  filter(restaurant == "Dairy Queen")


```

## Exercise 1

```{r}
#| title: plots

fat_calories <- fastfood |> 
  filter( restaurant %in% c("Mcdonalds", "Dairy Queen"))

ggplot( fat_calories, aes(x = cal_fat)) +
  geom_histogram( bins = 30, fill = "steelblue", color = "white")

```

```{r}
#| title: normal-dist


dqmean <- mean(dairy_queen$cal_fat)
dqsd   <- sd(dairy_queen$cal_fat)

ggplot(data = dairy_queen, aes(x = cal_fat)) +
  geom_blank() +
  geom_histogram(aes(y = ..density..)) +
  stat_function(fun = dnorm, args = c(mean = dqmean, sd = dqsd), col = "tomato")

```

## Exercise 2

Based on the line the data does seem to follow a close to normal distribution.

```{r}
#| title: evaluation


ggplot(data = dairy_queen, aes(sample = cal_fat)) + 
  geom_line(stat = "qq") +
  stat_qq_line()


sim_norm <- rnorm(n = nrow(dairy_queen), mean = dqmean, sd = dqsd)
```

## Exercise 3

```{r}
#| title: qqtest

ggplot( data = NULL, aes(sample = sim_norm))+
  geom_line(stat = "qq") +
  stat_qq_line()

```

```{r}
#| title: cont

qqnormsim(sample = cal_fat, data = dairy_queen)

```

## Exercise 4

Yes, the simulation looks similar to the plot created for the simulated data and brings us closer to showing that the the calories follow a normal distribution.

## Exercise 5

```{r}
#| title: mcdonalds-dist

qqnormsim( sample = cal_fat, data = mcdonalds)

```

# Normal Probabilities 

```{r}
#| title: normals 

1 - pnorm(q = 600, mean = dqmean, sd = dqsd)

dairy_queen %>% 
  filter(cal_fat > 600) %>%
  summarise(percent = n() / nrow(dairy_queen))

```

## Exercise 6

Two questions we can ask are what is the probability that a Taco Bell item is over 400 calories and what is the probability that an item at burger king is under 300 calories

```{r}
#| title: prob-question

taco <- fastfood %>% filter(restaurant == "Taco Bell")
bk   <- fastfood %>% filter(restaurant == "Burger King")

tb_mean <- mean(taco$calories)
tb_sd   <- sd(taco$calories)

q1_theoretical    <- 1 - pnorm(400, mean = tb_mean, sd = tb_sd)
q1_empirical <- mean(taco$calories > 400)   

bk_mean <- mean(bk$cal_fat)
bk_sd   <- sd(bk$cal_fat)

q2_theoretical   <- pnorm(300, mean = bk_mean, sd = bk_sd)
q2_empirical <- mean(bk$cal_fat < 300)
```

The theoretical probability for taco bell is `r  q1_theoretical` while the emperical is `r q1_empirical` compared to the burger king theoretical and emperical number being `r q2_theoretical` ,`r q2_empirical` respectively. The burger king probability is much closer to the theoretical probability

## Exercise 7 

```{r}
#| title: normals

ggplot(fastfood, aes(sample = sodium)) +
  geom_point(stat = "qq") +
  stat_qq_line() +
  facet_wrap(~ restaurant, scales = "free_y") +
  labs(title = "Normal probability plots of sodium by restaurant",
       x = "Theoretical quantiles", y = "Sodium (mg)")

```

it is tough to tell which restaurant is the closest to normally distributed for sodium content, but it looks like Chick Fil-A is the nearest to the line with just two larger outliers. Ultimately, it may be Burger King that is the closest to normally distributed.

## Exercise 8

These distributions appear to be stepwise due to a uniform amount of salt being used in the recipe. Each restaurant is likely to use a number like 250 mgs as a unit for each increase in salt.

## Exercise 9 

```{r}
#| title: carbs

ggplot(data = mcdonalds, aes(sample = total_carb)) +
  geom_point(stat = "qq") +
  stat_qq_line()

ggplot(data = mcdonalds, aes(x = total_carb)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "white")

```

It looks like the carbohydrate of mcdonalds is right skewed with most of the clustering being on the left side of the histogram.
